---
title: "Stats 2 Project 1"
author: "Michael Burgess"
date: "5/31/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
#data import
life_exp <- read.csv(choose.files(), header = T, sep = ",")

colSums(is.na(life_exp))

missing_data <- life_exp[rowSums(is.na(life_exp)) > 0,]
write.csv(missing_data, choose.files())

data_2015 <- life_exp[life_exp$Year == 2015,]
colSums(is.na(data_2015))

dim(data_2015)

#drop all the missing data so we can look at correlation and possibly make predicitions for missing values
not_missing <- life_exp %>% drop_na()

#convert character columns to factors
not_missing[sapply(not_missing, is.character)] <- lapply(not_missing[sapply(not_missing, is.character)], 
                                       as.factor)

#remove the first 3 columns since these don't contain predictors
pairs_data <- not_missing[,-c(1:3)]

#plot relationship between variables
pairs(pairs_data, lower.panel=NULL)
```

## 177 of 183 countries in 2015 are missing Alcohol data

## infant deaths and deaths under 5 redundant
## thinness 1-19 and 5-9 are redundant
## adult mortality and hiv.aids appear correlated
## gdp and percent expendatre are correlated
## polio and diptheria may be coorelated 
## maybe some linear relationship between alcohol and schooling
## bmi may be correlated with alchol and income composition
## income composition and schooling strong correlation
## strong linear correlation between life exptancy and income and 
## schooling
## BMI is broken almost 150 at nearly 60 BMI

```{r}
#using data without NAs here trying to get something clean so we can go back and #predict NA values for the main dataset

#remove redundant columns
non_redundant <- pairs_data[,-c(2,3)]

#remove BMI due to errors
non_redundant <- non_redundant[,-c(6)]

#remove non-sensical outliers, i.e. values over 1000 for deaths per 1000 people

non_redundant <- non_redundant %>% filter(under.five.deaths <= 1000)
non_redundant <- non_redundant %>% filter(Measles <= 1000)
non_redundant <- non_redundant %>% filter(HIV.AIDS <= 1000)

pairs(non_redundant, lower.panel = NULL)
```